---
title: "word2vec model trained on google news dataset"
subtitle: 'Determining the most similar words for proper nouns (MSBA classmate names)'
author: "Shubhankar Tiwari"
date: "`r Sys.Date()`"
abstract: |
  It is very hard for any machine learning algorithm to develop connections between proper nouns - Names, Places, Things. Most text processing techniques (like LDA) can group words like love, affection, care into a cluster/ topic with high probability because they observe them occuring frequently in even a small corpus  (like a novel). This is however very hard to do for nouns like names of people. Deep learning comes to the rescue. Google has trained a word2vec model on the massive google news dataset and has opened it for general use and experimentation. I used the model to analyse if it could establish any relations between the names of my classmates. I only used their first names to ensure privacy. The results are fantastic! 
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Loading Libraries

```{r}
library(knitr)
opts_chunk$set(echo=TRUE,cache=TRUE)

suppressPackageStartupMessages({
  library(readxl) #reading excel files
  library(dplyr) 
  library(visNetwork) #creating visual network

})

```


## Loading files

```{r}
nodes = read_excel('D:/Google Drive/Spring 2018/Courses/LSTM/word2vec/fuzzyRaw.xlsx',
                   sheet = "nodes",col_names = TRUE)
edges = read_excel('D:/Google Drive/Spring 2018/Courses/LSTM/word2vec/fuzzyRaw.xlsx',
                   sheet = "edges",col_names = TRUE)
```

## Network Graph

```{r}
edges$arrows = "to"

edges$color = ifelse(edges$group=="India",'#009688',
                       ifelse(edges$group=="Murica","#3F51B5",
                              ifelse(edges$group=="China","#9C27B0","#FFC107")))


nodes$title <- nodes$id
nodes$label <- nodes$id

visNetwork(nodes,edges, main = 'Connections between MSBA students',
           width = "1000px",height = "800px") %>%
  visLayout(randomSeed = 40) %>%
  visOptions(highlightNearest = list(enabled = TRUE, hover = TRUE),
             selectedBy = "group",nodesIdSelection = TRUE) %>%
  visGroups(groupname=as.vector("India"),shape="icon",
            icon=list(code="f111",color="#009688"))%>%
  visGroups(groupname=as.vector("Murica"),shape="icon",
            icon=list(code="f0c8",color="#3F51B5"))%>%
  visGroups(groupname=as.vector("China"),shape="icon",
            icon=list(code="f192",color="#9C27B0"))%>%
  addFontAwesome() %>%
  visLegend(width = 0.05) %>%
  visIgraphLayout(layout = "layout_on_sphere") %>%
  visNodes(size = 5)
```




## Insights

1. The model is clearly able to separate names based on nationality. We observe that there are no connections between people of different nationalities^[However American citizens with Indian origin have been classified as being related to Indian names]

2. Common names like `Gaurav` and `Saurabh` have a lot of connections

3. The model is also able to make a distinction in terms of gender for all the nationalities. Albeit names like Saumya (used by both genders) are ambiguous and the model cannot make any distinction.

4. Overall it's very easy for a human to connect `Cory` with `Kyle` and `Nikhil` with `Mayank` but it's very hard to train an algorithm to recognize this specially since such a training data set is hard to come by


The `edges` table has no connections b/w students of a different nationality. All the data has been preprocessed^[The python script is in my github repo] in python and excel before importing into R


```{r, echo=FALSE, results='asis'}
knitr::kable(head(edges, 10))
```

## Stay Tuned

Please visit my [github page](http://shubhstiws.github.io) for more such analysis. Comments, bug reports and
pull requests are always welcome.

## References

1. [Data](https://carlsonschool.umn.edu/degrees/master-science-in-business-analytics): Carlson School of Management, University of Minnesota, MS in Business Analytics, class of 2018

2. [Visual Network documentation](https://www.htmlwidgets.org/)

3. Google's pre trained word2vec [model](https://drive.google.com/file/d/0B7XkCwpI5KDYNlNUTTlSS21pQmM/edit). It's 1.5 GB

4. Word2vec [tutorial](http://mccormickml.com/2016/04/19/word2vec-tutorial-the-skip-gram-model/) by Chris McCormick

5. [Python notebook]() - preprocessing the data
